Relict deciduous oaks show high resilience to drought despite the effect of land-use legacies

### 2019-04-02 man/figures/bai_cronos.Rmd
udate nls

lms <- data.frame(year = as.numeric(c(2018, 2018, 2018)),
                     site = c("CA-Low", "CA-High", "SJ"), 
                     r2 = as.numeric(c(r2caL, r2caH, r2SJ)),
                     pvalue = as.numeric(round(c(pCaL, pCaH, pSJ), 4)),
                     y = as.numeric(c(as.numeric(predict(mCaL)[40]), 
                           as.numeric(predict(mCaH)[40]), 
                           as.numeric(predict(mSJ)[40]))))
lms$y <- lms$y / 100
lms <- lms %>% mutate(
  pc = ifelse(pvalue < 0.001, "p < 0.001", as.character(paste0("p = ", pvalue))),
  label = round(r2,3))

@@

# cores 
cronos_bai <- cro_sites %>% filter(nseries > 5) %>% 
  ggplot(aes(x=year, y=bai_mean/100, colour=site)) + 
  theme_bw() + ylab('BAI') + 
  geom_ribbon(aes(ymin = (bai_mean - bai_se)/100,
                  ymax = (bai_mean + bai_se)/100,
                  fill=site), alpha=.4, colour=NA) +
  geom_line() + 
  geom_point(size=.8) + 
  scale_colour_manual(values = colores) +
  scale_fill_manual(values = colores) +
  ylab(expression(paste("BAI (", cm^2, year^-1, ")", sep=''))) +
  scale_x_continuous(breaks = seq(1815, 2020, by=10), limits=c(1815,2020), sec.axis = dup_axis()) +
  geom_smooth(data = subset(cro_sites, year > 1975 & site == 'CA-High'), 
              aes(x=year, y=bai_mean/100, colour=site), method = 'lm', se=FALSE, size = .5) +
  geom_smooth(data = subset(cro_sites, year > 1975 & site == 'CA-Low'), 
              aes(x=year, y=bai_mean/100, colour=site), method = 'lm', se=FALSE, size = .5) +
  geom_smooth(data = subset(cro_sites, year > 1975 & site == 'SJ'), 
              aes(x=year, y=bai_mean/100, colour=site), method = 'lm', se=FALSE, size = .5)  +
  geom_text_repel(data = lms, 
                  aes(x=year, y = y, colour = site, label = label),
                  nudge_x = Inf, 
                  direction = "x", show.legend = FALSE, 
                  size = 3) + 
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(size=6),
        legend.position = c(0.2,.8),
        strip.background = element_blank(),
        axis.title.x.top = element_blank(), 
        axis.text.x.top = element_blank())  
```

@@

**Figure 4.** Basal Area Increment (BAI) chronologies of *Q. pyrenaica* for northern population (SJ; *orange*) and southern ones: low-elevation (CA-Low; *blue*) and high-elevation (CA-High, *black*) sites. Shading areas correspond to standard error of the mean. Number of series are displayed in the upper plot. We only show years replicated with # series > 5. Linear trends since 1975 are shown all sites (numbers indicate $r^2$ values).

>>>>

lms <- data.frame(year = as.numeric(c(2018, 2018, 2018)),
                     site = c("CA-Low", "CA-High", "SJ"), 
                     r2 = as.numeric(c(r2caL, r2caH, r2SJ)),
                     pvalue = as.numeric(round(c(pCaL, pCaH, pSJ), 4)),
                     y = as.numeric(
                       c(as.numeric(predict(mCaL)[40]), 
                           as.numeric(predict(mCaH)[40]), 
                           as.numeric(predict(mSJ)[40]))))
# lms$y <- lms$y / 100
lms <- lms %>% mutate(
  y = y/100,
  pc = ifelse(pvalue < 0.001, "p < 0.001", as.character(paste0("p = ", pvalue))),
  label = round(r2,3), 
  labelp = ifelse(pvalue < 0.001, paste0(label, "**"), label))

@@

# cores 
cronos_bai <- cro_sites %>% filter(nseries > 5) %>% 
  ggplot(aes(x=year, y=bai_mean/100, colour=site)) + 
  theme_bw() + ylab('BAI') + 
  geom_ribbon(aes(ymin = (bai_mean - bai_se)/100,
                  ymax = (bai_mean + bai_se)/100,
                  fill=site), alpha=.4, colour=NA) +
  geom_line() + 
  geom_point(size=.8) + 
  scale_colour_manual(values = colores) +
  scale_fill_manual(values = colores) +
  ylab(expression(paste("BAI (", cm^2, year^-1, ")", sep=''))) +
  scale_x_continuous(breaks = seq(1815, 2020, by=10), limits=c(1815,2020), sec.axis = dup_axis()) +
  geom_smooth(data = subset(cro_sites, year > 1975 & site == 'CA-High'), 
              aes(x=year, y=bai_mean/100, colour=site), method = 'lm', se=FALSE, size = .5) +
  geom_smooth(data = subset(cro_sites, year > 1975 & site == 'CA-Low'), 
              aes(x=year, y=bai_mean/100, colour=site), method = 'lm', se=FALSE, size = .5) +
  geom_smooth(data = subset(cro_sites, year > 1975 & site == 'SJ'), 
              aes(x=year, y=bai_mean/100, colour=site), method = 'lm', se=FALSE, size = .5)  +
  geom_text_repel(data = lms, 
                  aes(x=year, y = y, colour = site, label = labelp),
                  direction = "x", nudge_x = Inf, nudge_y = 1, show.legend = FALSE, 
                  size = 3) + 
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(size=6),
        legend.position = c(0.2,.8),
        strip.background = element_blank(),
        axis.title.x.top = element_blank(), 
        axis.text.x.top = element_blank())  
```

@@

**Figure 4.** Basal Area Increment (BAI) chronologies of *Q. pyrenaica* for northern population (SJ; *orange*) and southern ones: low-elevation (CA-Low; *blue*) and high-elevation (CA-High, *black*) sites. Shading areas correspond to standard error of the mean. Number of series are displayed in the upper plot. We only show years replicated with # series > 5. Linear trends since 1975 are shown all sites (numbers indicate $r^2$ values; asterisks indicate significant linear trend, *p < 0.001*).

### 2019-04-02 man/figures/resilience_long_term.Rmd
udate nls

>>>>

```{r nls}
# subset de datos para Non linear squares
d <- sdf %>% 
  dplyr::select(d_severity, rs_mean_mean) %>% 
  rename(sev = d_severity, rs = rs_mean_mean) %>% as.data.frame()

library("nlstools")

f <- as.formula(rs ~ a-b*exp(-c*sev))

preview(f, data=d, start=list(a=-1.6, b=-0.2, c=0.4))

# install nls2
devtools::install_github("ggrothendieck/nls2")

fo <- rs ~ Const + B * (sev ^ A)
st1 <- expand.grid(Const = seq(-2, 2, len = 10),
B = seq(-2,2, len = 10), A = seq(-1, 1, len = 10))
mod1 <- nls2(fo, start = st1, data=d, algorithm = "brute-force")
mod1
# use nls object mod1 just calculated as starting value for
# nls optimization. Same as: nls(fo, start = coef(mod1))
nls2(fo, start = mod1)

summary(mod1)

plot(sdf$d_severity, sdf$rs_mean_mean, pch=19)
lines(d$sev, predict(mod, list(sev = d$sev)))

### 2019-04-02 man/ms.Rmd
udate nls

The strength of the response of the resilience indices was related to drought severity. The exploration of resilience metrics of tree-growth for all severe droughts in our climatic data `REVIEW`(1900 - 2017), revealed a positive relation between resilience (*Rs*) and drought severity ($r^2$ = 0.336) (Figure 6). A similar pattern was found for recovery (*Rc*; $r^2$ = 0.438) but negative for resistance (*Rt*, $r^2$ = 0.155). Resilience values of tree-growth for 2005 was the lowest of the drought events analyzed (Figure 6) even not being the most severe drought events. `REVIEW` (CONEXION con el siguiente parrafo)

@@

The exploration of resilience metrics of tree-growth for all severe droughts in our climatic data (period 1900-2017?),

>>>>

At long-term perspective, the strength of the response of the resilience indices was related to drought severity. The exploration of resilience metrics of tree-growth for all severe droughts in our climatic data (1900 - 2017) (Table S3), revealed a positive relation between resilience (*Rs*) and drought severity ($r^2$ = 0.336) (Figure 6). A similar pattern was found for recovery (*Rc*; $r^2$ = 0.438) but negative for resistance (*Rt*, $r^2$ = 0.155).

Similar patterns were obtanied Resilience values (*Rs*) varied for the two last severe drought events (2005, 2012; Table 3). We obtained significantly higher values for the 2012, the most severe event (Table S3), than for 2005 in both variables, namely greenness and tree-growth (Tables S1-S2; Figure 7). Differences were more pronunced for BAI between drought event than for EVI (Tables S1-S2)

Resilience values of tree-growth for 2005 was the lowest of the drought events analyzed (Figure 6) even not being the most severe drought events. `REVIEW` (CONEXION con el siguiente parrafo)

@@
